In this section we consider the unweighted variants of the problem.
We describe and analyze a local search algorithm, 
and show that when the maximum possible capacity, $C$, is fixed, the algorithm achieves 
a $\frac{C + 1}{2C}$ approximation ratio.
Throughout this section we consider $C$ to be fixed, and $k$ to be a fixed parameter.  

The local search algorithm maintains a feasible matching throughout its execution
and operates in 
iterative manner where in each iteration it tries to find a better solution by
replacing a subset of the edges in the current solution with 
another (larger) subset of edges not in the solution.
The local search algorithm halts when no improvement can be done.
The algorithm is described in Algorithm~\ref{alg:local} 

\begin{algorithm}
\SetKw{True}{true}
\SetKw{False}{false}
\KwIn{$G = (V, E)$, $c : V \rightarrow \N$, $k$}
\KwOut{$M$}
$M \leftarrow \emptyset$								\\
\Repeat{done}{
\label{line:outerloop}
 	$done \leftarrow{}$ \True							\\
 	\For{$M' \in M : |M'| \leq Ck$}{
 		\For{$V' \in V \setminus M : |V'| = |M'| + 1$}{
			\If{$M \setminus M' \cup V'$ is feasible}{
				$M \leftarrow{} M \setminus M' \cup V'$	\\
				$done \leftarrow{}$ \False				\\
			}
		}
 	}
}
\Return{$M$}

\caption{
\label{alg:local}
Local Search}
\end{algorithm}


We claim that the local search algorithm halts in polynomial time.
Observed that in every iteration the algorithm either improves the value of the solution
by one or otherwise it terminates. 
Thus, after a maximum of $n$ iterations the algorithm stops.
In every iteration we examine all the subsets of edges of a fixed size and test for feasibility,
both these operations can be done in polynomial time. 

To analyze the approximation ratio achieved by the local search algorithm consider an 
arbitrary but fixed optimal matching $M^* = \{s_1, \ldots, s_m\}$,
where $s_1, \ldots, s_m$ are the stars induced by the optimal solution.
Given the solution outputted by the local search algorithm, 
define the following graph $L = (\{v_i : s_i \in M^* \}, E)$ 
where $E = \{(v_i, v_j) : \exists (u,v) \in M \land u \in s_i, v \in s_j \}$.
That is, each vertex in $L$ corresponds to a star in $M^*$ and there is an edge 
between two vertices if and only if there is an edge $(u,v)$ in $M$ connecting the two 
corresponded stars.
Denote by $d^*_i$ the sum of the degrees in $s_i$, and by $d_i$ the sum of the degrees of 
edges from $M$ intersected with $s_i$. \todo{better define this last line}
\todo{add an illustration}
